---
title: "In-class Exercise 3"
author: "Georgia Ng"
date: "September 2, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## 3.1 Overview

## 3.2 Installing the Required Packages

Since maptools is retired and binary is removed from CRAN, we will be downloading it from the posit public package manager snapshots.

> Note: It is important to add `eval:false` in the code chunk as shown below after installation is complete to avoid it being executed every time the quarto document is being rendered.

```{#| eval: false}
install.packages("maptools",repos = "https://packagemanager.posit.co/cran/2023-10-13")
```

```{r}
pacman::p_load(sf,tmap,tidyverse)
```

## 3.3 The Data

```{r}
mpsz_sf <- st_read(dsn="data/MasterPlan2014SubzoneBoundaryWebSHP/", 
                   layer="MP14_SUBZONE_WEB_PL")
```

The code chunk below, `st_union()` is used to derive the **coastal outline** sf tibble data.frame.

```{r}
sg_sf <-mpsz_sf %>%
  st_union()
```

```{r}
plot(sg_sf)
```

## 3.4 Viewing

The below chunk of code imports the ACLED Myanmar data, converts it into a spatial format, changes the coordinate system, and formats the event dates into a standard date format.

```{r}
acled_sf <- read_csv("data/ACLED_Myanmar.csv") %>% 
  st_as_sf(coords = c(
    "longitude", "latitude"), crs = 4326) %>% 
  st_transform(crs= 32647)%>%
  mutate(event_date = dmy(event_date))
```

This code produces an interactive map displaying dots for events in 2023 or classified as "Political violence.

```{r}
#| eval: false
tmap_mode('view')
acled_sf %>%
  filter(year == 2023 |
           event_type == "Political violence") %>%
  tm_shape()+
  tm_dots()
tmap_mode('plot')
```
