---
title: "In-class Exercise 4: Spatio-Temporal Point Patterns Analysis"
author: "Georgia Ng"
date: "September 10, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
  cache: true
---

## 4.0 Overview

A spatio-temporal point process (also called space-time or spatial-temporal point process) is a random collection of points, where each point represents the time and location of an event. Examples of events include incidence of disease, sightings or births of a species, or the occurrences of fires, earthquakes, lightning strikes, tsunamis, or volcanic eruptions. In this lesson, you will learn the basic concepts and methods of Spatio-temporal Point Patterns Analysis. You will also gain hands-on experience on using these methods to discover real-world point processes.

The specific questions we would like to answer are:

-   are the locations of forest fire in Kepulauan Bangka Belitung spatial and spatio-temporally independent?

-   if the answer is NO, where and when the observed forest fire locations tend to cluster?

## 4.1 Importing the Packages

For the purpose of this study, five R packages will be used. They are:

-   `rgdal` for importing geospatial data in GIS file format such as shapefile into R and save them as Spatial\*DataFrame,

-   `maptools` for converting Spatial\* object into ppp object,

-   `raster` for handling raster data in R,

-   `sparr` provides functions to estimate fixed and adoptive kernel-smoothed spatial relative risk surfaces via the density-ratio method and perform subsequent inference. Fixed-bandwidth spatiotemporal density and relative risk estimation is also supported.

-   `spatstat` for performing Spatial Point Patterns Analysis such as kcross, Lcross, etc., and

-   `tmap` for producing cartographic quality thematic maps.

```{r}
pacman::p_load(sf, raster, spatstat, sparr, tmap, tidyverse)
```

## 4.2 The Data

For the purpose of this exercise, two data sets will be used, they are:

-   *forestfires*, a csv file provides locations of forest fire detected from the Moderate Resolution Imaging Spectroradiometer (MODIS) sensor data. The data are downloaded from [Fire Information for Resource Management System](https://firms.modaps.eosdis.nasa.gov/download/). For the purpose of this exercise, only forest fires within Kepulauan Bangka Belitung will be used.

-   *Kepulauan_Bangka_Belitung*, an ESRI shapefile showing the sub-district (i.e. [kelurahan](https://en.wikipedia.org/wiki/Villages_of_Indonesia)) boundary of Kepulauan Bangka Belitung. The data set was downloaded from [Indonesia Geospatial](https://www.indonesia-geospasial.com/2023/05/download-shapefile-batas-administrasi.html) portal. The original data covers the whole Indonesia. For the purpose of this exercise, only sub-districts within Kepulauan Bangka Belitung are extracted.

### 4.2.1 Importing Study Area

```{r}
kbb <- st_read(dsn="data/rawdata/", 
                   layer="Kepulauan_Bangka_Belitung")
```

::: callout-note
Notice that uniquely the polygon in the geometry column when imported is of a **Polygon Z** type. This means that each polygon not only defines a 2D shape but also includes elevation data with a z-coordinate. This additional z-dimension allows for a more detailed representation of the polygon's geometry, incorporating vertical information such as elevation or depth.
:::

The below revised code chunk serves to do the following:

1.  Group the boundaries up
2.  Drop the Z values
3.  Transform the coordinate system

```{r}
kbb_sf <- st_read(dsn="data/rawdata/", 
                   layer="Kepulauan_Bangka_Belitung") %>%
  st_union()%>%
  st_zm(drop = TRUE, what = "ZM")%>%
  st_transform(32748)

```

### 4.2.2 Converting OWIN

Next, as.owin() is used to convert kbb into an owin object, which is a spatial window or region of interest for point pattern analysis. Once converted to an `owin` object, we can use it with functions from spatial point pattern analysis packages, such as `spatstat`, to analyze point patterns within the defined boundary. It helps in setting up the spatial context for further analysis.

```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```

Next, `class()` is used to confirm if the output is indeed an owin object.

```{r}
class(kbb_owin)

```

### 4.2.3 Importing and Preparing Forest Fire Data

Next, we will import the forest fire data set into the R environment. The code reads forest fire data from a CSV file, converts it into an `sf` object using longitude and latitude coordinates, and then reprojects the spatial data from WGS84 to the UTM zone 48S coordinate system. This prepares the data for further spatial analysis in a projection appropriate for the region of interest.

```{r}
fire_sf <- read_csv("data/rawdata/forestfires.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"),
           crs=4326)%>%
  st_transform(crs = 32748)
```

Because `ppp` object only accept numeric or character as mark. The code chunk below is used to convert data type of `acq_date` to numeric.

```{r}
fire_sf <- fire_sf %>%
  mutate(DayofYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date, label= TRUE, abbr = FALSE))
```

## 4.3 Visualising The Fire Points

### 4.3.1 Overall Plot

```{r}
tm_shape(kbb_sf)+
  tm_polygons() +
tm_shape(fire_sf) +
  tm_dots()
```

### 4.3.2 Visualising Geographic Distribution Of Forest Fires By Month

```{r}
tm_shape(kbb_sf) +
  tm_polygons()+
  tm_shape(fire_sf)+
  tm_dots(size= 0.1)+
  tm_facets(by="Month_fac", free.coords = FALSE, drop.units= TRUE)
```

## 4.4 Computing STKDE by Month

### 4.4.1 Extracting forest fires by month

The code chunk below is used to remove the unwanted fields from `fire_sf` sf data.frame. This is because `as.ppp()` only need the mark field and geometry field from the input sf data frame.

```{r}
fire_month <- fire_sf %>% 
  select(Month_num)
```

### 4.4.2 Creating ppp

The code chunk below is used to derive a ppp object called fire_month from fire_month of data frame.

```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```

The code chunk below is used to check the output is in the correct object class.

```{r}
summary(fire_month_ppp)
```

Next, we will check if there are duplicated point events by using the code chunk below.

```{r}
any(duplicated(fire_month_ppp))
```

### 4.4.3 Including Owin Object

The code chunk below is used to combine origin_am_ppp and am_owin objects into one.

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```

```{r}
plot(fire_month_owin)
```

### 4.4.4 Computing Spatio-Temporal KDE

Next, `spattemp.density()` of sparr package is used to compute the `STKDE`.

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

### 4.4.5 Plotting the spatio-temporal KDE object

In the code chunk below, `plot()` of R base is used to the KDE for between July 2023 - December 2023.

```{r}
tims <- c(7,8,9,10,11,12)
par(mfcol=c(2,3))
for(i in tims){ 
  plot(st_kde, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at month",i))
}
```

## 4.5 Computing STKDE by Day of Year

In this section, you will learn how to computer the STKDE of forest fires by day of year.

### 4.5.1 Creating ppp object

In the code chunk below, DayofYear field is included in the output ppp object.

```{r}
fire_yday_ppp <- fire_sf %>% 
  select(DayofYear) %>%
  as.ppp()
```

### 4.5.2 Including Owin object

Next, code chunk below is used to combine the ppp object and the owin object.

```{r}
fire_yday_owin <- fire_yday_ppp[kbb_owin]
summary(fire_yday_owin)
```

## 4.5.3

```{r}
kde_yday <- spattemp.density(
  fire_yday_owin)
summary(kde_yday)
```

```{r}
#| eval: false
plot(kde_yday)
```

## 4.6 Computing STKDE by Day of Year: Improved method

One of the nice function provides in sparr package is `BOOT.spattemp()`. It support bandwidth selection for standalone spatiotemporal density/intensity based on bootstrap estimation of the MISE, providing an isotropic scalar spatial bandwidth and a scalar temporal bandwidth.

Code chunk below uses `BOOT.spattemp()` to determine both the spatial bandwidth and the scalar temporal bandwidth.

```{r}
set.seed(1234)
BOOT.spattemp(fire_yday_owin) 
```

### 4.6.1 Computing spatio-temporal KDE

Now, the STKDE will be derived by using h and lambda values derive in previous step.

```{r}
kde_yday <- spattemp.density(
  fire_yday_owin,
  h = 9000,
  lambda = 19)
summary(kde_yday)

```

### 4.6.2 Plotting the output spatio-temporal KDE

Last, `plot()` of **sparr** package is used to plot the output as shown below.

```{r}
#| eval: false
plot(kde_yday)
```
